---
title: "CS 109B: Final Project - Milestone 2: "
subtitle: "April 6, 2017"
output: html_document
---

```{r,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(109) # Set seed for random number generator
#load packages
library(ggplot2)
library(dplyr)
library(cluster)
library(factoextra)
library(mclust)
library(corrplot)
library(dbscan)
library(MASS)
library(ggfortify)
library(NbClust)
library(caret)
library('e1071')
library(tidyr)
```

# Question: What does your choice of Y look like?

In IMDB data set, we have total 28 genres. However, it has been observared that some of these gernes show a high correlation. For instance, Thriller movie trends to associate closely with Horror movie. Also, some genres have very small number of movies, using clustering method helps us in reducing the number of genres. In specific, we intend to perform clustering on closely related genres (based on avialable attributes), so that we will end up with grouping together movie genres which show high similarity based on movie attributes. After clustering, we will have certain number of clusters, then we assign the top combination of genres in each cluster as the response variable for that cluster. For example, if the most popular multiple-label genre for cluster1 is Romance, Drama and Action, then we will regard the movies in cluster 1 have genre label: Romance, Drama and Action.  

# Load Data

```{r}
# load data
data.df <- read.csv("C:/Users/cheriexu/Downloads/CS109B/imdb_imputed_byRF_for_cluster_with_response.csv")

#since most columns are categorical (dummy coding will be too many columns), 
#we used the relative frequency for later clustering
relative_req <- function(x){
  if(class(x) == "numeric"){
    tab <- table(x)/nrow(data.val)
    ind <- match(x, as.numeric(names(tab)))
    unname(tab[ind])
  }
  else{
    x <- as.character(x)
    tab <- table(x)/nrow(data.val)
    ind <- match(x, as.character(names(tab)))
    unname(tab[ind])
  }
}
```

# Clustering

Considering that we will have more than 200,000 movies in the whole data set, we would like to choose clustering methods that are suitable for very large data sets. (The following three clustering method were covered in the lectures, so we won't explain them in details here).

K-means: able to handle very large data set; good for general-purpose; not very uneven cluster size; will not have too many clusters.

PAM: less suitable for very large data set; minimize the average dissimilarity of objects to their closest selected object.

DBSCAN: able to handle very large data set; usually generate uneven cluster sizes

We would like to compare the performance of clustering based on their silhouette plot and whether the number of observations in clusters is reasonable. 

```{r}
#convert all categorical variables to relative frequency
data.val <- data.df %>% dplyr::select(c(1:32)) %>% mutate('certificate' = ifelse(certificates_R == 1, "R", ifelse(certificates_PG == 1, "PG","None")))
data.val[,'certificate'] <- relative_req(data.val[,'certificate'])
data.val[,4:ncol(data.val)] <- scale(data.val[,4:ncol(data.val)], center = F)
data.scaled <- data.val[,4:ncol(data.val)]
head(data.scaled)
```

**Kmeans** 
```{r}
#choose optimal number of clusters

# elbow plots
fviz_nbclust(data.scaled, kmeans,iter.max=30, method="wss", nstart = 5) + 
  ggtitle("K-means clustering  - optimal number of clusters") +
  geom_vline(xintercept=5, linetype=2)

# average silhouette widths
fviz_nbclust(data.scaled, kmeans,method="silhouette", nstart = 5) +
  ggtitle("K-means clustering  - optimal number of clusters") 

``` 

Since the average silhouette widths give too few clusters, we will like to choose the optimal number of clusters basde on the elbow plots.

```{r}
cluster <- 5
data.km <- kmeans(data.scaled, cluster, nstart = 5)

```

```{r}
# silhouette plot
sil_kmeans <- silhouette(data.km$cluster, dist = daisy(data.scaled))
fviz_silhouette(sil_kmeans) + ggtitle("silhouette plot for the kmeans clustering")

```


 **Partitioning around medoids** (PAM) 

```{r}
# PAM
# find optimal cluster

# elbow plots
fviz_nbclust(data.scaled, pam, method="wss") + 
  ggtitle("PAM clustering - optimal number of clusters") +
  geom_vline(xintercept=6,linetype=2)

# average silhouette widths
fviz_nbclust(data.scaled,pam,method="silhouette") +
  ggtitle("PAM clustering - optimal number of clusters") 

```

```{r}
cluster <- 6
data.pam = pam(data.scaled, k=cluster)

```

```{r}
# silhouette plot
fviz_silhouette(silhouette(data.pam), 
  main="Silhouette plot for PAM clustering")

```


 Density-based clustering

```{r}
kNNdistplot(data.scaled,k=3)
abline(3.0,0,lty=2,lwd=2,col="red") # added after seeing kNN plot
#knees around 3.0

data.db = dbscan(data.scaled, eps=3, minPts = 3)

data.db.df <- as.data.frame(data.db$cluster)
table(data.db.df)
colnames(data.db.df)[1] <- "cluster"

#outliers
data.db.df %>%
  filter(cluster == 0)
```


# Clustering comparison:

Since we would like to compare the performance of clustering based on their silhouette plot and whether the number of observations in clusters is reasonable, K-means out-performs PAM based on silhouette plot, and K-means gives much more reasonable number of observations in each cluster. Although the cluster distributin in K-means is still imbalanced (as we would expected), it is in a reasonable range. However, for the result we get from DBSCAN, we usually ends up with one cluster with too many observations than a reasonable expection (even choosing different parameters). Thus, we would like use K-means as the final clustering methods to generate response variables.


# Process response variable from clustering result:

We proposed two methods to process response variables

(1) In each cluster, compute the occurance of each individual genre, then choose top three genres from them. Combine the top three genres from each cluster as the new label for that cluster. For example, the top three populat genres in cluster 1 are Romance, Drama and Action, then the new label for cluster 1 is Romance, Drama, Action, together as the respones variable. However, this processing has two drawbacks. The first one is that it is very likely more than one cluster having the same top three genres. Also, some genres are more popular than others in the whole data set, so it will be more likely to be in the top three genres in each cluster.

(2) Thus, we proposed an alternative solution, by choosing the top one genre combination from each cluster as the response variable. It is very unlikely to have the same genre combination for different clusters using this method, and it will be consistent with the results based on the clustering. The main goal of clustering is to merge some similar genres into others, by using this method, we can satisify our inital goal. 

```{r}
#(1) compute each cluster's top three genres 
data.df <- data.df %>% dplyr::mutate(cluster_response = data.km$'cluster')
for (i in 1:length(unique(data.df$cluster_response))){
  df1 <- data.df %>% filter(cluster_response == i) %>% 
    dplyr::select(c(Sci.Fi:Biography))
  sort(colSums(df1))
}

genres <- colnames(data.df)[33:56]
genre_list = c("Sci.Fi","Crime", "Romance", "Animation", "Music",
"Adult", "Comedy", "War", "Horror", "Film.Noir", "Western", "Thriller",
"Adventure","Mystery","Short", "Drama", "Action", "Documentary", "Musical","History","Family","Fantasy","Sport","Biography")
data.df$genrecomb <- do.call(paste0, data.df[genre_list])

#(2) compute each cluster's top one genre combination
for (i in 1:length(unique(data.df$cluster_response))){
  df2 <- data.df %>% tidyr::unite(col =genres, Sci.Fi:Biography, sep = ",")
  df3 <- df2 %>% filter(cluster_response == i)
  print(genre_list[strsplit(df2[which.max(relative_req(df3$genres)),'genres'],",")[[1]]=="1"])
}

```

For our sample data, we end up with five clusters, and we figure out the genre combination for each cluster as following. Then we assign the result to the original data set.

```{r}
map <-data.frame(cluster_response = c(1:5), genres_comb = c("Horror, Thriller, Drama",
                                                            "Romance, Music, Drama",
                                                            "Animation, Comedy, Adventure, Family, Fantasy",
                                                            "Romance, Comedy, Fantasy",
                                                            "Horror, Thriller, Action"))
```

Output the original data set with new response variable as a csv file
```{r}
df5 <- left_join(data.df %>% dplyr::select(-ncol(data.df)), map, by = "cluster_response" )
write.csv(x = df5, file = "imdb_cluster_result.csv",row.names = F)
```

